---
title: "BuildSdmTMB"
author: "Stephanie Hopkins"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    
---

```{r setup, include=FALSE}

# Run first if not already installed
# remotes::install_github("pbs-assess/sdmTMBextra", dependencies = TRUE)
 

rm(list=ls())
gc()

load.lib <- c("tidyverse", "data.table", "cowplot", "sf", "rcompanion",
              "pROC", "here", "sdmTMB", "sdmTMBextra", "visreg",
              "ggforce", # enables plot_anisotropy
              "spatstat",# Get summary table for nearest neighbor
              "rsample", "timetk", "zoo") 

install.lib <- load.lib[!load.lib %in% installed.packages()]

for(lib in install.lib) install.packages(lib,dependencies=TRUE)
sapply(load.lib,require,character=TRUE)

sf_use_s2(FALSE)

dir <- here(getwd())

```

# Load YearSeason data

```{r, load.seasonal.data}

# Training set 2
train2 <- fread(here(dir, "Data", "ExampleRun", "train2"))

# Test set 2
test2 <- fread(here(dir, "Data", "ExampleRun", "test2"))

# YearSeason gaps to fill
load(file = here(dir, "Data", "ExampleRun",
                                    "est.missing.time2.rda"))
  
```

# Construct YearSeason mesh: 

see https://pbs-assess.github.io/sdmTMB/articles/basic-intro.html: cutoff is in the units of X and Y (km here), represents minimum distance between knots before a new mesh vertex is added). Muhling et al., 2025 tried a bunch of cutoff values, and chose one that 1) had good out-of-sample predictability, and 2) wasn't too "blocky". They found Values between ~ 50 and 200 were fairly reasonable. 

Important: cannot use anisotropy with barrier mesh (see warning when run sdmTMB).

```{r, construct.mesh}

# Load land barrier
land.barrier <- read_sf(here(dir, "Data", "Shapefiles", "land.barrier.shp"))

# Cutoff values control the minimum distance between knots (here in km).
## It is generally better to start with a coarser mesh (larger cutoff)
## However there is a tradeoff on spatial predictability (more knots) and
## over fitting the time to process. If the day is irregularly distributed
## you can also try residual-based knot placement
meshTrain2 <- make_mesh(train2, xy_cols = c("X", "Y"), cutoff = 50) 

# proj_scaling should match units of (since we are working in m, but density
# is in km, divide by 1000)
barrier2 <- add_barrier_mesh(meshTrain2, land.barrier, proj_scaling = 1000)

## If greater than >1,000–2,000 nodes, 
## you’re likely in trouble unless you have a lot of RAM.
length(barrier2$mesh$loc[,1])
# [1] 604

# Example plot code from ?add_barrier_mesh
mesh_df_water2 <- barrier2$mesh_sf[barrier2$normal_triangles, ]
mesh_df_land2 <- barrier2$mesh_sf[barrier2$barrier_triangles, ]

Mesh.with.barrier2 <-
ggplot() +
  geom_sf(data = land.barrier) +
  geom_sf(data = mesh_df_water2, size = 1, colour = "blue") +
  geom_sf(data = mesh_df_land2, size = 1, colour = "green")

Mesh.with.barrier2

ggsave2(filename=paste("Seasonal.Mesh.with.barrier.jpeg",sep=''),
        plot=Mesh.with.barrier2, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)


plot(barrier2)
points(train2$X, train2$Y, col = "red", pch = 19, 
       cex = 0.2)

```

# Run YearSeason models

## Try with spatial effects

```{r, make.seasonal.sptl.model.w.fix.tim}

set.seed(2) 

## Add age and length fixed effects with temporal and spatial autocorrelation
## Note that this is testing presence absence and then density predicted by age
## and length bin so we want the intercept
fit10 <- sdmTMB(Density ~ as.factor(age) + as.factor(len_bin) +
                 as.factor(YearSeas.dup),
               data = train2, 
               mesh = barrier2,
               # Fit model by year and month
               ## Note if time is specified, spatiotemporal correlation is tested
               ## by default
               time = "YearSeas.nm",
               # Use "ar1" or "rw" if you expect temporal autocorrelation
               # Otherwise, use default "iid" (independent by time step)
               spatiotemporal = "ar1",
               spatial = "on", 
               extra_time = est.missing.time2,
               # Fit hurdle model (first link is 0 or 1, second link is > 0)
               family = delta_gamma(link1 = "logit", link2 = "log"),
               anisotropy = FALSE,
               # Show progress
               silent = FALSE)

# Check model skill, fit, and convergence
sanity(fit10)
# ✔ Non-linear minimizer suggests successful convergence
# ✔ Hessian matrix is positive definite
# ✔ No extreme or very small eigenvalues detected
# ✔ No gradients with respect to fixed effects are >= 0.001
# ✔ No fixed-effect standard errors are NA
# ✔ No standard errors look unreasonably large
# ✔ No sigma parameters are < 0.01
##------------------------------------------------------------------------------
## ✖ `sigma_E` is larger than 100
## ℹ Consider simplifying the model or adding priors
##
#### Large values of sigma_E indicate that there is a high level of spatial 
#### variation that your model has not explained. 
##------------------------------------------------------------------------------
# ✔ Range parameters don't look unreasonably large

summary(fit10)
## M1:
#### Spatiotemporal AR1 correlation (rho): 0.08
#### Matérn range: 82.73
#### Spatial SD: 81.67
#### Spatiotemporal marginal AR1 SD:  625.07

## M2:
#### Dispersion Parameter: 0.45 (Good, typical range: 0.5-1.5)
#### Spatiotemporal AR1 Correlation (rho): 0.44 (Moderate to strong, 
## typical range: 0.4 - 0.8)
#### Matérn Range: 119.278 (Good, typical range: 50-200 depending on scale)
#### Spatial SD: 10.11
#### Spatiotemporal Marginal AR1 SD: 47.38 (Good, typical range: 10-50)
#### ML Criterion: 4289.604 (Best used for model comparison, not absolute assessment)

# Predict (gives a whole new df, unlike GAM)
pTrain10 <- predict(fit10, newdata = train10, type = "response")

pTest10 <- predict(fit10, newdata = test10, type = "response")

# Overall R2 for test and train 
summary(lm(Density ~ est, data = pTrain10))$r.squared # 0.140585
summary(lm(Density ~ est, data = pTest10))$r.squared # 2.217959e-06

# MAE
mean(abs(pTrain10$Density - pTrain10$est)) # 314.6031
mean(abs(pTest10$Density - pTest10$est)) # 0.9113865

mean(train2$Density)
# [1] 231.8522
sd(train2$Density)
# [1] 11672.67

mean(test2$Density)
# [1] 0.9113857                                                               
sd(test2$Density)
# [1] 24.58594

fit10_op <- run_extra_optimization(fit10, newton_loops = 10)
max(fit10$gradients)
max(fit10_op$gradients)

sanity(fit10_op)

```

## Try with AR1 and Age specific Intercept

```{r, age.treated.as.independant.ar1}

set.seed(2) 

# Add age and length fixed effects with temporal and spatial autocorrelation
# Note that this is testing presence absence and then density predicted by age
## and length bin so we want the intercept
# The 0 intercept indicates that age should be treated independently
## with there own intercepts
fit11 <- sdmTMB(Density ~ 0 + as.factor(age) + as.factor(len_bin),
               data = train2, 
               mesh = barrier2,
               # Fit model by year and month
               time = "YearSeas.nm",
               # Use "ar1" or "rw" if you expect temporal autocorrelation
               # "ar1" assumes density is stationary and reverts to the mean
               # "rw" assumes "non-stationary"
               # Otherwise, use default "iid" (independent per time step)
               spatiotemporal = "ar1",
               spatial = "off", 
               extra_time = est.missing.time2,
               # Fit hurdle model (first link is 0 or 1, second link is > 0)
               family = delta_gamma(link1 = "logit", link2 = "log"),
               anisotropy = FALSE,
               silent = TRUE)

# Check model skill, fit, and convergence
sanity(fit11)
# ✔ Non-linear minimizer suggests successful convergence
# ✔ Hessian matrix is positive definite
# ✔ No extreme or very small eigenvalues detected
## ✖ `ln_kappa` gradient > 0.001
## ℹ See ?run_extra_optimization(), standardize covariates, and/or simplify the model
# ✔ No fixed-effect standard errors are NA
# ✔ No standard errors look unreasonably large
# ✔ No sigma parameters are < 0.01
## ✖ `sigma_E` is larger than 100
## ℹ Consider simplifying the model or adding priors
# ✔ Range parameters don't look unreasonably large

summary(fit11)
# M1:
## Spatiotemporal AR1 correlation (rho): 0.10
## Matérn range: 77.17
## Spatiotemporal marginal AR1 SD: 594.86
# M2: 
## Dispersion parameter: 0.45
## Spatiotemporal AR1 correlation (rho): 0.48
## Matérn range: 121.11
## Spatiotemporal marginal AR1 SD: 49.19
## ML criterion at convergence: 4290.250

# Predict (gives a whole new df, unlike GAM)
pTrain11 <- predict(fit11, newdata = train2, type = "response") 
pTest11 <- predict(fit11, newdata = test2, type = "response")

# Overall R2 for test and train 
summary(lm(Density ~ est, data = pTrain11))$r.squared # 0.1411138
summary(lm(Density ~ est, data = pTest11))$r.squared # 8.335531e-05

# MAE
mean(abs(pTrain11$Density - pTrain11$est)) # 314.9381
mean(abs(pTest11$Density - pTest11$est)) # 0.9113859

mean(train2$Density)
# [1] 231.8522
sd(train2$Density)
# [1] 11672.67

mean(test2$Density)
# [1] 0.9113857                                                               
sd(test2$Density)
# [1] 24.58594 

fit11_op <- run_extra_optimization(fit11, newton_loops = 10)
max(fit11$gradients)
max(fit11_op$gradients)

sanity(fit11_op)

``` 

```{r, age.treated.as.independant.ar1}

set.seed(2) 

# Add age and length fixed effects with temporal and spatial autocorrelation
# Note that this is testing presence absence and then density predicted by age
## and length bin so we want the intercept
# The 0 intercept indicates that age should be treated independently
## with there own intercepts
fit12 <- sdmTMB(Density ~ 0 + as.factor(age),
               data = train2, 
               mesh = barrier2,
               # Fit model by year and month
               time = "YearSeas.nm",
               # Use "ar1" or "rw" if you expect temporal autocorrelation
               # "ar1" assumes density is stationary and reverts to the mean
               # "rw" assumes "non-stationary"
               # Otherwise, use default "iid" (independent per time step)
               spatiotemporal = "ar1",
               spatial = "off", 
               extra_time = est.missing.time2,
               # Fit hurdle model (first link is 0 or 1, second link is > 0)
               family = delta_gamma(link1 = "logit", link2 = "log"),
               anisotropy = FALSE,
               silent = TRUE)

# Check model skill, fit, and convergence
sanity(fit12)
# ✔ Non-linear minimizer suggests successful convergence
# ✔ Hessian matrix is positive definite
# ✔ No extreme or very small eigenvalues detected
# ✔ No gradients with respect to fixed effects are >= 0.001
# ✔ No fixed-effect standard errors are NA
# ✔ No standard errors look unreasonably large
# ✔ No sigma parameters are < 0.01
## ✖ `sigma_E` is larger than 100
## ℹ Consider simplifying the model or adding priors
# ✔ Range parameters don't look unreasonably large

summary(fit12)
# M1:
## Spatiotemporal AR1 correlation (rho): 0.10
## Matérn range: 76.43
## Spatiotemporal marginal AR1 SD: 592.59
# M2: Dispersion parameter: 0.42
## Spatiotemporal AR1 correlation (rho): 0.48
## Matérn range: 125.45
## Spatiotemporal marginal AR1 SD: 50.77
## ML criterion at convergence: 4403.838

# Predict (gives a whole new df, unlike GAM)
pTrain12 <- predict(fit12, newdata = train2, type = "response") 
pTest12 <- predict(fit12, newdata = test2, type = "response")

# Overall R2 for test and train 
summary(lm(Density ~ est, data = pTrain12))$r.squared # 0.06580561
summary(lm(Density ~ est, data = pTest12))$r.squared # 2.459787e-06

# MAE
mean(abs(pTrain12$Density - pTrain12$est)) # 358.7532
mean(abs(pTest12$Density - pTest12$est)) # 0.9113859

mean(train2$Density)
# [1] 231.8522
sd(train2$Density)
# [1] 11672.67

mean(test2$Density)
# [1] 0.9113857                                                               
sd(test2$Density)
# [1] 24.58594

```

## Try with AR1 and No Age specific Intercept

#```{r, age.treated.as.dependant.ar1}

set.seed(2) 

# Add age fixed effects with temporal and spatial autocorrelation
# Note that this is testing presence absence and then density predicted by age
## and length bin so we want the intercept
fit13 <- sdmTMB(Density ~ as.factor(age),
               data = train2, 
               mesh = barrier2,
               # Fit model by year and month
               ## Note if time is specified, spatiotemporal correlation is tested
               ## by default
               time = "YearSeas.nm",
               # Use "ar1" or "rw" if you expect temporal autocorrelation
               ## "ar1" assumes density is stationary and reverts to the mean
               ## "rw" assumes "non-stationary"
               ## Otherwise, use default "iid" (independent by time step)
               spatiotemporal = "ar1",
               spatial = "off", 
               extra_time = est.missing.time2,
               # Fit hurdle model (first link is 0 or 1, second link is > 0)
               family = delta_gamma(link1 = "logit", link2 = "log"),
               anisotropy = FALSE,
               # Show progress
               silent = FALSE)

# Check model skill, fit, and convergence
sanity(fit13)
# ✔ Non-linear minimizer suggests successful convergence
# ✔ Hessian matrix is positive definite
# ✔ No extreme or very small eigenvalues detected
# ✔ No gradients with respect to fixed effects are >= 0.001
# ✔ No fixed-effect standard errors are NA
# ✔ No standard errors look unreasonably large
# ✔ No sigma parameters are < 0.01
## ✖ `sigma_E` is larger than 100
## ℹ Consider simplifying the model or adding priors
# ✔ Range parameters don't look unreasonably large
# Predict (gives a whole new df, unlike GAM)

summary(fit13)
# M1: 
## Spatiotemporal AR1 correlation (rho): 0.10
## Matérn range: 76.43
## Spatiotemporal marginal AR1 SD: 592.59
# M2: 
## Dispersion parameter: 0.42
## Spatiotemporal AR1 correlation (rho): 0.48
## Matérn range: 125.45
## Spatiotemporal marginal AR1 SD: 50.77
## ML criterion at convergence: 4403.838

pTrain13 <- predict(fit13, newdata = train2, type = "response")
pTest13 <- predict(fit13, newdata = test2, type = "response")

# Overall R2 for test and train 
summary(lm(Density ~ est, data = pTrain13))$r.squared # 0.06580561
summary(lm(Density ~ est, data = pTest13))$r.squared # 2.459787e-06

# MAE
mean(abs(pTrain13$Density - pTrain13$est)) # 358.7532 
mean(abs(pTest13$Density - pTest13$est)) # 0.9113859

mean(train2$Density)
# [1] 231.8522
sd(train2$Density)
# [1] 11672.67

mean(test2$Density)
# [1] 0.9113857                                                               
sd(test2$Density)
# [1] 24.58594

#```

## Try with Random Walk and Age specific Intercept

```{r, age.treated.as.independant.rw}

set.seed(2) 

# Add age and length fixed effects with temporal and spatial autocorrelation
# Note that this is testing presence absence and then density predicted by age
## and length bin so we want the intercept
# The 0 intercept indicates that age and length bin should be treated independently
## with there own intercepts
fit14 <- sdmTMB(Density ~ 0 + as.factor(age),
               data = train2, 
               mesh = barrier2,
               # Fit model by year and month
               ## Note if time is specified, spatiotemporal correlation is tested
               ## by default
               time = "YearSeas.nm",
               # Use "ar1" or "rw" if you expect temporal autocorrelation
               ## "ar1" assumes density is stationary and reverts to the mean
               ## "rw" assumes "non-stationary"
               ## Otherwise, use default "iid" (independent by time step)
               spatiotemporal = "rw",
               spatial = "off", 
               extra_time = est.missing.time2,
               # Fit hurdle model (first link is 0 or 1, second link is > 0)
               family = delta_gamma(link1 = "logit", link2 = "log"),
               anisotropy = FALSE,
               # Show progress
               silent = FALSE)

# Check model skill, fit, and convergence
sanity(fit14)
# ✔ Non-linear minimizer suggests successful convergence
# ✔ Hessian matrix is positive definite
# ✔ No extreme or very small eigenvalues detected
# ✔ No gradients with respect to fixed effects are >= 0.001
# ✔ No fixed-effect standard errors are NA
# ✔ No standard errors look unreasonably large
# ✔ No sigma parameters are < 0.01
## ✖ `sigma_E` is larger than 100
## ℹ Consider simplifying the model or adding priors

summary(fit14)
# M1:
## Matérn range: 294.71
## Spatial SD: 0.01
## Spatiotemporal RW SD: 1902.36
#M2:
## Dispersion parameter: 0.40
## Matérn range: 261.23
## Spatial SD: 66.07
## Spatiotemporal RW SD: 51.69
## ML criterion at convergence: 4591.333

# Predict (gives a whole new df, unlike GAM)
pTrain14 <- predict(fit14, newdata = train2, type = "response")
pTest14 <- predict(fit14, newdata = test2, type = "response")

# Overall R2 for test and train 
summary(lm(Density ~ est, data = pTrain14))$r.squared # 0.1010444
summary(lm(Density ~ est, data = pTest14))$r.squared # 5.544298e-07

# MAE
mean(abs(pTrain14$Density - pTrain14$est)) # 332.5696
mean(abs(pTest14$Density - pTest14$est)) # 2.421233

mean(train2$Density)
# [1] 231.8522
sd(train2$Density)
# [1] 11672.67

mean(test2$Density)
# [1] 0.9113857                                                               
sd(test2$Density)
# [1] 24.58594

```

## Try with Random Walk and No Age specific Intercept

```{r, age.treated.as.dependant.rw}

set.seed(2)

# Add age and length fixed effects with temporal and spatial autocorrelation
# Note that this is testing presence absence and then density predicted by age
## and length bin so we want the intercept
fit15 <- sdmTMB(Density ~ as.factor(age),
               data = train2, 
               mesh = barrier2,
               # Fit model by year and quarter
               time = "YearSeas.nm",
               # Use "ar1" or "rw" if you expect temporal autocorrelation
               ## "ar1" assumes density is stationary and reverts to the mean
               ## "rw" assumes "non-stationary"
               ## Otherwise, use default "iid" (independent by time step)
               spatiotemporal = "rw",
               spatial = "on", 
               extra_time = est.missing.time2,
               # Fit hurdle model (first link is 0 or 1, second link is > 0)
               family = delta_gamma(link1 = "logit", link2 = "log"),
               anisotropy = FALSE)

# Check model skill, fit, and convergence
sanity(fit15)
# ✔ Non-linear minimizer suggests successful convergence
# ✔ Hessian matrix is positive definite
# ✔ No extreme or very small eigenvalues detected
# ✔ No gradients with respect to fixed effects are >= 0.001
# ✔ No fixed-effect standard errors are NA
# ✔ No standard errors look unreasonably large
# ✔ No sigma parameters are < 0.01
## ✖ `sigma_E` is larger than 100
## ℹ Consider simplifying the model or adding priors
# ✔ Range parameters don't look unreasonably large

summary(fit15)
# M1:
## Matérn range: 294.71
## Spatial SD: 11.66
## Spatiotemporal RW SD: 1902.36
# M2:
## Dispersion parameter: 0.40
## Matérn range: 261.23
## Spatial SD: 66.07
## Spatiotemporal RW SD: 51.69
## ML criterion at convergence: 4591.333

# Predict (gives a whole new df, unlike GAM)
pTrain15 <- predict(fit15, newdata = train2, type = "response")
pTest15 <- predict(fit15, newdata = test2, type = "response")

# Overall R2 for test and train 
summary(lm(Density ~ est, data = pTrain15))$r.squared # 0.08367316
summary(lm(Density ~ est, data = pTest15))$r.squared # 1.870584e-06

# MAE
mean(abs(pTrain15$Density - pTrain15$est)) # 332.5696
mean(abs(pTest15$Density - pTest15$est)) # 2.421233 

mean(train2$Density)
# [1] 231.8522
sd(train2$Density)
# [1] 11672.67

mean(test2$Density)
# [1] 0.9113857                                                               
sd(test2$Density)
# [1] 24.58594

```


## Try with Random Walk and Age specific Intercept

```{r, age.treated.as.dependant.rw}

set.seed(2)

# Add age and length fixed effects with temporal and spatial autocorrelation
# Note that this is testing presence absence and then density predicted by age
## and length bin so we want the intercept
fit16 <- sdmTMB(Density ~ 0 + as.factor(age),
               data = train2, 
               mesh = barrier2,
               # Fit model by year and quarter
               time = "YearSeas.nm",
               # Use "ar1" or "rw" if you expect temporal autocorrelation
               ## "ar1" assumes density is stationary and reverts to the mean
               ## "rw" assumes "non-stationary"
               ## Otherwise, use default "iid" (independent by time step)
               spatiotemporal = "rw",
               spatial = "off", 
               extra_time = est.missing.time2,
               # Fit hurdle model (first link is 0 or 1, second link is > 0)
               family = delta_gamma(link1 = "logit", link2 = "log"),
               anisotropy = FALSE)

# Check model skill, fit, and convergence
sanity(fit16)
# ✔ Non-linear minimizer suggests successful convergence
# ✔ Hessian matrix is positive definite
# ✔ No extreme or very small eigenvalues detected
# ✔ No gradients with respect to fixed effects are >= 0.001
# ✔ No fixed-effect standard errors are NA
# ✔ No standard errors look unreasonably large
# ✔ No sigma parameters are < 0.01
## ✖ `sigma_E` is larger than 100
## ℹ Consider simplifying the model or adding priors
# ✔ Range parameters don't look unreasonably large

summary(fit16)
# M1:
## Matérn range: 294.71
## Spatiotemporal RW SD: 1902.36
# M2:
## Dispersion parameter: 0.40
## Matérn range: 258.70
## Spatiotemporal RW SD: 53.69
## ML criterion at convergence: 4592.622

# Predict (gives a whole new df, unlike GAM)
pTrain16 <- predict(fit16, newdata = train2, type = "response")
pTest16 <- predict(fit16, newdata = test2, type = "response")

# Overall R2 for test and train 
summary(lm(Density ~ est, data = pTrain16))$r.squared # 0.08396251
summary(lm(Density ~ est, data = pTest16))$r.squared # 2.370368e-06

# MAE
mean(abs(pTrain16$Density - pTrain16$est)) # 332.0232
mean(abs(pTest16$Density - pTest16$est)) # 3.077845 

mean(train2$Density)
# [1] 231.8522
sd(train2$Density)
# [1] 11672.67

mean(test2$Density)
# [1] 0.9113857                                                               
sd(test2$Density)
# [1] 24.58594

```
```{r}

BIC(fit11)
BIC(fit12)
BIC(fit13)
BIC(fit14)
BIC(fit15)
BIC(fit16)

# Model 12 and 13 are better, so focus on the one with independent age coefficients 
# Extract residuals from the binomial (zero-inflation) component
set.seed(123)
pTrain12$M1Resids <- residuals(fit12, type = "mle-mvn", model = 1)
hist(pTrain12$M1Resids[is.finite(pTrain12$M1Resids)]) 
qqnorm(pTrain12$M1Resids[is.finite(pTrain12$M1Resids)]);abline(0, 1)

ggplot(subset(pTrain12, is.finite(M1Resids)), aes(X, Y, col = M1Resids)) + 
  scale_colour_viridis_c() +
  geom_point() + 
  facet_wrap(~YearSeas.nm) + 
  coord_fixed()

set.seed(123)
# Leave M2Resids as-is
pTrain12$M2Resids <- residuals(fit12, type = "mle-mvn", model = 2)
# Plot histogram/QQ-plot with finite values only
hist(pTrain12$M2Resids[is.finite(pTrain12$M2Resids)])
qqnorm(pTrain12$M2Resids[is.finite(pTrain12$M2Resids)]); abline(0, 1)

# Plot map only with finite residuals
ggplot(subset(pTrain12, is.finite(M2Resids)), aes(X, Y, col = M2Resids)) +
  scale_colour_viridis_c() +
  geom_point() +
  facet_wrap(~YearSeas.nm) +
  coord_fixed()


```

