
---
title: "BuildSdmTMB"
author: "Stephanie Hopkins"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    
---

```{r setup, include=FALSE}

# Run first if not already installed
# remotes::install_github("pbs-assess/sdmTMBextra", dependencies = TRUE)
# remotes::install_github("pbs-assess/sdmTMB")

rm(list=ls())
gc()

load.lib <- c("tidyverse", "data.table", "cowplot", "sf", "rcompanion",
              "pROC", "here", "sdmTMB", "sdmTMBextra", "visreg",
              "ggforce", # enables plot_anisotropy
              "spatstat",# Get summary table for nearest neighbor
              "rsample", "timetk", "zoo") 

install.lib <- load.lib[!load.lib %in% installed.packages()]

for(lib in install.lib) install.packages(lib,dependencies=TRUE)
sapply(load.lib,require,character=TRUE)

sf_use_s2(FALSE)

dir <- here(getwd())

```

# Load previous sdm data

Build new adult sardine SDM using environmental predictors from the 3km WC15 ROMS.
Vars are SST, surface chl-a, upper 50m zooplankton biomass, SSB. 
Chl-a didn't improve AUC, dropped for now.

```{r, load.previous.dat}

# Where previous data is stored
sdmDir <- here(dir, "Data", "Previous.sdmTMB.Data")

# Load data
dat <- readRDS(paste(sdmDir, "/sardineTrainingData.rds", sep = ""))

# Project data
dat2 <- dat %>%
  filter(yr > 2002 & survey == 'cps') %>% 
  distinct() %>%
  group_by(yr) %>%
  mutate(hauls = n_distinct(haul)) %>%
  ungroup() %>% 
  select(yr, cruise, haul, lat, hauls) %>% 
  distinct()
# View(dat2)

test2003 <- dat %>% filter(yr == 2003 & survey == 'cps')

length(unique(test2003$haul))
# [1] 41
# View(test2003)

```

# Load new sdm data

There will likely be some hauls not previously accounted, but the number of hauls should be similar. Use 2003 as a test run.

```{r, load.new.dat}

areaTranBuffsum.polygons <- 
   read_sf(here(dir, "Data", "Shapefiles", "cluster.with.area.defs.shp"))
#View(areaTranBuff.polygons)

unique(areaTranBuffsum.polygons$Area_id)
# [1] 2 3 4 5 1

sort(unique(areaTranBuffsum.polygons$Area_id))
# [1] 1 2 3 4 5

areaTranBuff.polygons.2003 <- areaTranBuffsum.polygons %>%
  filter(Year == 2003)

colnames(areaTranBuffsum.polygons)

```

# Create density estimates to be used for biomass abundance indices

Note that here we want to include the absence data and are not concerned by lengths

```{r, biomass.density.area2}

area2poly.dens <- areaTranBuffsum.polygons %>%
  mutate(Cruise = substr(keys,1,6)) %>%
  group_by(Cruise, Year, Month, Day, tim_stp, clstGRP, Hauls, NbClust, Area_id,
           Area_km2, geometry, age, len_bin) %>%
  # Note that some clusters will have both presence and absence data
  # but this will give you the proportion to apply in stock synthesis for length
  # and age
  summarise(WghtLenBin = sum(WghtLenBin), .groups = "drop") %>%
  mutate(
    # Calculate density
    Density = WghtLenBin / Area_km2,

    # Create proper date object
    Date = as.Date(paste(Year, Month, Day, sep = "-"), format = "%Y-%m-%d"),
    
    # Convert to numeric
    Year = as.numeric(Year),
    Month = as.numeric(Month),
    Day = as.numeric(Day),

    # Create YearQuarter
    YearQuarter = as.yearqtr(Date),

    # Extract numeric quarter
    Quarter = case_when(
      str_detect(YearQuarter, "Q1") ~ 1,
      str_detect(YearQuarter, "Q2") ~ 2,
      str_detect(YearQuarter, "Q3") ~ 3,
      str_detect(YearQuarter, "Q4") ~ 4
    ),

    # Create numeric time step index from baseline year (e.g., 2003)
    NumericQuarter = (Year - min(Year)) * 4 + Quarter,

  ) %>%
  select(-tim_stp)

# View(area2poly.dens)

plotNormalHistogram(area2poly.dens$Density)
range(area2poly.dens$Density)
# [1] 0.000 912149.7

```

# Extract polygon centers

```{r, area2.mesh.prestep}

# Extract polygon centers
area2poly.dens2 <- area2poly.dens %>%
  mutate(X = st_coordinates(st_centroid(.))[,1]/1000,
         Y = st_coordinates(st_centroid(.))[,2]/1000) %>%
  st_drop_geometry()

```

# Restructure data to include P/A

First find mixed clusters where both presence and absence values exist. Filter out clusters that have only absence values.

```{r, presence.absence.col.pt1}

# Find steps where there are absence values within a cluster
Absent.dat <- area2poly.dens2 %>%
  filter(is.na(age)) %>%
  select(-c(age,len_bin)) %>%
  # Add key to identify clusters
  mutate(vals = paste0(Cruise, Date, clstGRP, Area_id, Area_km2, X, Y))

dim(Absent.dat)
# [1] 1282   18

# Find steps where there are presence values within a cluster
Present.dat <- area2poly.dens2 %>%
  filter(!is.na(age)) %>%
  # Add presence binary value
  mutate(PA = 1) %>%
  # Add key to identify clusters
  mutate(vals = paste0(Cruise, Date, clstGRP, Area_id, Area_km2, X, Y))

# Clusters with both presence and absence should be given a 1 indicating presence, 
# and are but are accounted for in the density measure. 
Absent.dat.mxd <- Absent.dat %>%
  filter(vals %in% Present.dat$vals) %>%
  # Add presence binary value
  mutate(PA = 1)

# Find steps where the entire clusters has no Sardines
# These should be given a 0 binary value indicating total absence.
Absent.dat.all <- Absent.dat %>%
  filter(!vals %in% Present.dat$vals) %>%
  # Add absence binary value
  mutate(PA = 0)

```

## For each bin in the presence data, split by age length bin

Combine survey clusters with total absence and bin data so that for each cruise, there total absence clusters are appended.

```{r, presence.absence.col.pt2}

# Check that all Cruise values in Absent.dat.all are present in Present.dat
unique(Absent.dat.all$Cruise) %in% unique(Present.dat$Cruise)
#  [1]  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE FALSE
# [14]  TRUE  TRUE  TRUE FALSE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE
# [27]  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE  TRUE

#-------------------------------------------------------------------------------
# Store cruises where there are no shared Cruises
#-------------------------------------------------------------------------------

Absent.matched.cruise <- Absent.dat.all %>%
  filter(Cruise %in% Present.dat$Cruise)

dim(Absent.matched.cruise)
# [1] 1020   19   

## Check if there is at least a shared Date
Absent.matched.time <- Absent.matched.cruise %>%
  filter(Date %in% Present.dat$Date)

dim(Absent.matched.time)
# [1] 181  19

## Check if there is at least a shared Year and Month
Absent.matched.yrmon <- Absent.matched.cruise %>%
  filter(!Date %in% Present.dat$Date &
           Year %in% Present.dat$Year &
           Month %in% Present.dat$Month)

dim(Absent.matched.yrmon)
# [1] 839  19

#-------------------------------------------------------------------------------
# Store cruises where there are no shared Cruises
#-------------------------------------------------------------------------------

Absent.ToTunique <- Absent.dat.all %>%
  filter(!Cruise %in% Present.dat$Cruise)

dim(Absent.ToTunique)
# [1] 44 19

## Check if there is at least a shared Date
Absent.ToTunique.date <- Absent.ToTunique %>%
  filter(Date %in% Present.dat$Date)

dim(Absent.ToTunique.date)
# [1] 24 19

## Check if there is at least a shared Year and Month
Absent.ToTunique.yrmon <- Absent.ToTunique %>%
  filter(!Date %in% Present.dat$Date &
           Year %in% Present.dat$Year &
           Month %in% Present.dat$Month)

dim(Absent.ToTunique.yrmon)
# [1] 20 19

```

## For shared Cruise and Date values, bind rows

```{r, presence.absence.col.pt3.1}

#-------------------------------------------------------------------------------
# For when a shared date exists
#-------------------------------------------------------------------------------

# Define keys
join_keys <- c("Cruise", "Date")

cruise.date.dat <- Present.dat %>%
  group_by(age, len_bin) %>%
  group_split() %>%
  map_df(~ {
    present_grp <- .x

    # Get rows from Absent.dat.all that match on Cruise
    absent_matches <- Absent.matched.time %>%
      filter(Cruise %in% Present.dat$Cruise & 
               Date %in% Present.dat$Date) %>%
      semi_join(present_grp, by = join_keys) %>%
      mutate(age = unique(present_grp$age),
         len_bin = unique(present_grp$len_bin))

    # Append matched Absent rows to this group
    bind_rows(present_grp, absent_matches)
  })

```

## For shared Cruise and Year Month values, bind rows

```{r, presence.absence.col.pt3.2}

# Define keys
join_keys <- c("Cruise", "Year", "Month")

cruise.yrmon.dat <- Present.dat %>%
  group_by(age, len_bin) %>%
  group_split() %>%
  map_df(~ {
    present_grp <- .x

    # Get rows from Absent.dat.all that match on Cruise
    absent_matches <- Absent.matched.yrmon %>%
      filter(Cruise %in% Present.dat$Cruise &
               !Date %in% Present.dat$Date &
               Year %in% Present.dat$Year &
               Month %in% Present.dat$Month) %>%
      semi_join(present_grp, by = join_keys) %>%
      mutate(age = unique(present_grp$age),
         len_bin = unique(present_grp$len_bin))

    # Append matched Absent rows to this group
    bind_rows(present_grp, absent_matches)
  })

```

## For non-shared Cruise data and Date only, bind rows

```{r, presence.absence.col.pt3.3}

# Define keys
join_keys <- c("Date")

date.dat.only <- Present.dat %>%
  group_by(age, len_bin) %>%
  group_split() %>%
  map_df(~ {
    present_grp <- .x

    # Get rows from Absent.dat.all that match on Cruise
    absent_matches <- Absent.ToTunique.date %>%
      filter(!Cruise %in% Present.dat$Cruise &
               Date %in% Present.dat$Date) %>%
      semi_join(present_grp, by = join_keys) %>%
      mutate(age = unique(present_grp$age),
         len_bin = unique(present_grp$len_bin))

    # Append matched Absent rows to this group
    bind_rows(present_grp, absent_matches)
  })

```

## For non-shared Cruise data and Year Month only, bind rows

```{r, presence.absence.col.pt3.4}

# Define keys
join_keys <- c("Year", "Month")

yrmon.dat.only <- Present.dat %>%
  group_by(age, len_bin) %>%
  group_split() %>%
  map_df(~ {
    present_grp <- .x

    # Get rows from Absent.dat.all that match on Cruise
    absent_matches <- Absent.ToTunique.yrmon %>%
      filter(!Cruise %in% Present.dat$Cruise &
               !Date %in% Present.dat$Date &
               Year %in% Present.dat$Year &
               Month %in% Present.dat$Month) %>%
      semi_join(present_grp, by = join_keys) %>%
      mutate(age = unique(present_grp$age),
         len_bin = unique(present_grp$len_bin))

    # Append matched Absent rows to this group
    bind_rows(present_grp, absent_matches)
  })

```

## Join back everything

```{r, join.full.data.presense.absence}

Full.dat <- rbind(cruise.date.dat, cruise.yrmon.dat, date.dat.only, yrmon.dat.only) %>%
  distinct() %>%
  # Reorder chronologically
  arrange(NumericQuarter)  %>%
  mutate(YearQuarter.dup = as.numeric(NumericQuarter))

summary(Full.dat)

unique(Full.dat$PA)

```

# Regroup Quarts as 2 Seasons

There is moderate to strong temporal autocorrelation in the model as the modelling mean absolute error is significantly reduced by inclusion of temporal effects. However the spatial effects are not be well captured.

```{r, 2season.structure}

TwoSeas.dat <- Full.dat %>%
  mutate(
    Season = case_when(
      month(Date) %in% 1:6 ~ "S1",
      month(Date) %in% 7:12 ~ "S2"
      ),
    # Create combined time step variable using year and season
    YearSeas = paste(year(Date), Season, sep = ""),
     Season.nm =  
      case_when(
      grepl("S1", Season) ~ 1,
      grepl("S2", Season) ~ 2
      ),
    # Combine year and numeric season
    YearSeas.nm = (Year + (Season.nm - 1) / 10),
    YearSeas.dup = as.numeric(YearSeas.nm))
# View(TwoSeas.dat)  

unique(TwoSeas.dat$YearSeas)
unique(TwoSeas.dat$YearSeas.nm)

```

# Find Missing Seasons

```{r, missing.seasons}

# Generate complete sequence
# Define the range of years
years <- 2003:2024

# Create the sequence of year-season combinations
year_season_sequence <- data.frame(
  YearSeas.nm = unlist(lapply(years, function(year) {
  # For each year, create two entries: season 1 (0.0) and season 2 (0.1)
  ## year + 0/10 gives 2003.0,while year + 1/10 gives 2003.1
  c(year + 0 / 10, year + 1 / 10)  
}))
)

# Print the resulting sequence
year_season_sequence

missing_seas <- year_season_sequence %>%
  anti_join(TwoSeas.dat, by = "YearSeas.nm") %>%
  select(YearSeas.nm)

missing_seas

```

# Create summaries

## For individuals less than 175 mm

```{r, length.bin.less.than.175mm}

less175 <- TwoSeas.dat %>%
  filter(len_bin %in% c("100", "125", "150", "175")) %>%
  # Drop age and collapse len_bin
  group_by(Cruise, Year, Month, Day, clstGRP, Hauls, NbClust, Area_id, Area_km2,
           Date, YearQuarter, Quarter, NumericQuarter, X, Y, PA, YearQuarter.dup,
           Season, YearSeas, Season.nm, YearSeas.nm, YearSeas.dup, vals) %>%
  summarise(WghtLenBin = sum(WghtLenBin), .groups = "keep") %>%
  mutate(lengths = "< 175 mm", 
         Density = WghtLenBin/Area_km2)

dim(less175)
# [1] 1059   26

dim(distinct(less175))
# [1] 1059   26

less175 <- less175 %>%
  mutate(Density.tr = log1p(Density))

# View(less175)

#-------------------------------------------------------------------------------
# Map density
#-------------------------------------------------------------------------------

# Base map
less175.den.map <- 
  ggplot() +
  geom_point(data = less175, aes(x=X,y=Y, col = Density)) +
  scale_color_viridis_c() +
  facet_wrap(~YearSeas) +
  title("Density by Year Season for lengths < 175 mm") +
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("less175.den.map.jpeg",sep=''),
        less175.den.map, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

# Log transformed map
less175.den.tr.map <- 
  ggplot() +
  geom_point(data = less175, aes(x=X,y=Y, col = Density.tr)) +
  scale_color_viridis_c() +
  facet_wrap(~YearSeas) +
  title("Density (ln(x+1)) by Year Season for lengths < 175 mm") +
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("less175.den.tr.map.jpeg",sep=''),
        less175.den.tr.map, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

#-------------------------------------------------------------------------------
# Separate positive densities
#-------------------------------------------------------------------------------

less175.pres <- less175 %>%
  filter(PA == 1) 

# Explore range (note 0.0 is an artefact of rounding precision)
range(less175.pres$Density)
# [1] 0.0 940504.5

#-------------------------------------------------------------------------------
# Histogram figs
#-------------------------------------------------------------------------------

# For base histogram
less175.pres.den.hist <-
  ggplot() +
  geom_histogram(data = less175.pres,
                 aes(x = Density, y = ..density..), 
                 bins = 50,  # You can adjust the number of bins or use binwidth
                 fill = "orange", color = "black", alpha = 0.7) + 
  facet_wrap(~YearSeas) +
  labs(title = "Histogram of Density by Year Season for lengths < 175 mm", 
       x = "Density", y = "Density (f(x))") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  # Clean, minimal theme
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("less175.pres.den.hist.jpeg",sep=''),
        less175.pres.den.hist, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

# For log transformed histogram
less175.pres.den.tr.hist <-
ggplot() +
  geom_histogram(data = less175.pres,
                 aes(x = Density.tr, y = ..density..), 
                 bins = 50,  # You can adjust the number of bins or use binwidth
                 fill = "orange", color = "black", alpha = 0.7) + 
  facet_wrap(~YearSeas) +
  labs(title = "Histogram of Density by Year Season for lengths < 175 mm", 
       x = "Density (ln(x+1))", y = "Density (f(ln(x+1)))") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  # Clean, minimal theme
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("less175.pres.den.tr.hist.jpeg",sep=''),
        less175.pres.den.tr.hist, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

#-------------------------------------------------------------------------------
# Boxplot figs
#-------------------------------------------------------------------------------

# For base boxplot
less175.pres.den.boxplot <-
  ggplot() +
  geom_boxplot(data = less175.pres, 
               aes(x = YearSeas, y = Density),
               fill = "orange", 
               color = "black", alpha = 0.7) +
  labs(title = "Density by Year Season for lengths < 175 mm", 
       x = "Year Season", y = "Density (x)") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 60, vjust = .5, hjust=.50),
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("less175.pres.den.boxplot.jpeg",sep=''),
        less175.pres.den.boxplot, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

# For log transformed boxplot
less175.pres.den.tr.boxplot <-
  ggplot() +
  geom_boxplot(data = less175.pres, 
               aes(x = YearSeas, y = Density.tr),
               fill = "orange", 
               color = "black", alpha = 0.7)  +
  labs(title = "Density by Year Season for lengths < 175 mm", 
       x = "Year Season", y = "Density (x)") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 60, vjust = .5, hjust=.50),
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("less175.pres.den.tr.boxplot.jpeg",sep=''),
        less175.pres.den.tr.boxplot, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

```

## For individuals greater than 175 mm

```{r, length.bin.greater.than.175mm}

greater175 <- TwoSeas.dat %>%
  filter(len_bin %in% c("200", "225", "250+")) %>%
  # Drop age and collapse len_bin
  group_by(Cruise, Year, Month, Day, clstGRP, Hauls, NbClust, Area_id, Area_km2,
           Date, YearQuarter, Quarter, NumericQuarter, X, Y, PA, YearQuarter.dup,
           Season, YearSeas, Season.nm, YearSeas.nm, YearSeas.dup, vals) %>%
  summarise(WghtLenBin = sum(WghtLenBin), .groups = "keep") %>%
  mutate(lengths = "> 175 mm", 
         Density = WghtLenBin/Area_km2)

dim(greater175)
# [1] 1283   26

dim(distinct(greater175))
# [1] 1283   26

greater175 <- greater175 %>%
  mutate(Density.tr = log1p(Density))

# View(greater175)

#-------------------------------------------------------------------------------
# Map density
#-------------------------------------------------------------------------------

# Base map
greater175.den.map <- 
  ggplot() +
  geom_point(data = greater175, aes(x=X,y=Y, col = Density)) +
  scale_color_viridis_c() +
  facet_wrap(~YearSeas) +
  title("Density by Year Season for lengths > 175 mm") +
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("greater175.den.map.jpeg",sep=''),
        greater175.den.map, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

# Log transformed map
greater175.den.tr.map <- 
  ggplot() +
  geom_point(data = greater175, aes(x=X,y=Y, col = Density.tr)) +
  scale_color_viridis_c() +
  facet_wrap(~YearSeas) +
  title("Density (ln(x+1)) by Year Season for lengths > 175 mm") +
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("greater175.den.tr.map.jpeg",sep=''),
        greater175.den.tr.map, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

#-------------------------------------------------------------------------------
# Separate positive densities
#-------------------------------------------------------------------------------

greater175.pres <- greater175 %>%
  filter(PA == 1) 

# Explore range (note 0.0 is an artefact of rounding precision)
range(greater175.pres$Density)
# [1] 0.0 940504.5


#-------------------------------------------------------------------------------
# Histogram figs
#-------------------------------------------------------------------------------

# For base histogram
greater175.pres.den.hist <-
  ggplot() +
  geom_histogram(data = greater175.pres,
                 aes(x = Density, y = ..density..), 
                 bins = 50,  # You can adjust the number of bins or use binwidth
                 fill = "lightgreen", color = "black", alpha = 0.7) + 
  facet_wrap(~YearSeas) +
  labs(title = "Histogram of Density by Year Season for lengths > 175 mm", 
       x = "Density", y = "Density (f(x))") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  # Clean, minimal theme
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("greater175.pres.den.hist.jpeg",sep=''),
        greater175.pres.den.hist, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

# For log transformed histogram
greater175.pres.den.tr.hist <-
ggplot() +
  geom_histogram(data = greater175.pres,
                 aes(x = Density.tr, y = ..density..), 
                 bins = 50,  # You can adjust the number of bins or use binwidth
                 fill = "lightgreen", color = "black", alpha = 0.7) + 
  facet_wrap(~YearSeas) +
  labs(title = "Histogram of Density by Year Season for lengths > 175 mm", 
       x = "Density (ln(x+1))", y = "Density (f(ln(x+1)))") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0)) +
  # Clean, minimal theme
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("greater175.pres.den.tr.hist.jpeg",sep=''),
        greater175.pres.den.tr.hist, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

#-------------------------------------------------------------------------------
# Boxplot figs
#-------------------------------------------------------------------------------

# For base boxplot
greater175.pres.den.boxplot <-
  ggplot() +
  geom_boxplot(data = greater175.pres, 
               aes(x = YearSeas, y = Density),
               fill = "lightgreen", 
               color = "black", alpha = 0.7) +
  labs(title = "Density by Year Season for lengths > 175 mm", 
       x = "Year Season", y = "Density (x)") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 60, vjust = .5, hjust=.50),
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("greater175.pres.den.boxplot.jpeg",sep=''),
        greater175.pres.den.boxplot, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

# For log transformed boxplot
greater175.pres.den.tr.boxplot <-
  ggplot() +
  geom_boxplot(data = greater175.pres, 
               aes(x = YearSeas, y = Density.tr),
               fill = "lightgreen", 
               color = "black", alpha = 0.7)  +
  labs(title = "Density by Year Season for lengths > 175 mm", 
       x = "Year Season", y = "Density (x)") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 60, vjust = .5, hjust=.50),
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("greater175.pres.den.tr.boxplot.jpeg",sep=''),
        greater175.pres.den.tr.boxplot, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

```

## For individuals by age

```{r, age.bin.only}

ageDens <- TwoSeas.dat %>%
  # Drop len_bin and collapse age
  group_by(Cruise, Year, Month, Day, clstGRP, Hauls, NbClust, Area_id, Area_km2,
           Date, YearQuarter, Quarter, NumericQuarter, X, Y, PA, YearQuarter.dup,
           Season, YearSeas, Season.nm, YearSeas.nm, YearSeas.dup, vals, age) %>%
  summarise(WghtLenBin = sum(WghtLenBin), .groups = "keep") %>%
  mutate(Density = WghtLenBin/Area_km2)

dim(ageDens)
# [1] 8200   26

dim(distinct(ageDens))
# [1] 8200   26

ageDens <- ageDens %>%
  mutate(Density.tr = log1p(Density))

# View(ageDens)

#-------------------------------------------------------------------------------
# Map density
#-------------------------------------------------------------------------------

# Base map
ageDens.den.map <- 
  ggplot() +
  geom_point(data = ageDens, aes(x=X,y=Y, col = Density)) +
  scale_color_viridis_c() +
  facet_grid(age ~ YearSeas) +
  # Clean, minimal theme
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("ageDens.den.map.jpeg",sep=''),
        ageDens.den.map, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

# Log transformed map
ageDens.den.tr.map <- 
  ggplot() +
  geom_point(data = ageDens, aes(x=X,y=Y, col = Density.tr)) +
  scale_color_viridis_c() +
  facet_grid(age ~ YearSeas) +
  # Clean, minimal theme
  theme_bw() +  
  # Customizing facet labels
  theme(strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10), 
        axis.title = element_text(size = 12, face = "bold"))

ggsave2(filename=paste("ageDens.den.tr.map.jpeg",sep=''),
        ageDens.den.tr.map, device="jpeg", 
        path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
        unit="cm", limitsize = FALSE)

#-------------------------------------------------------------------------------
# Separate positive densities
#-------------------------------------------------------------------------------

ageDens.pres <- ageDens %>%
  filter(PA == 1) 

# Explore range (note 0.0 is an artefact of rounding precision)
range(ageDens.pres$Density)
# [1] 0.0 940504.5


#-------------------------------------------------------------------------------
# Figures
#-------------------------------------------------------------------------------

for (a in unique(ageDens.pres$age)) {
  ageDens.pres.set <- ageDens.pres[ageDens.pres$age == a, ]
  age = unique(ageDens.pres.set$age)
  print(age)
  
  #-----------------------------------------------------------------------------
  # Map density
  #-----------------------------------------------------------------------------

  # Base map
  ageDens.den.map <- 
    ggplot() +
    geom_point(data = ageDens, aes(x=X,y=Y, col = Density)) +
    scale_color_viridis_c() +
    facet_wrap(~ YearSeas) +
    labs(title = paste("Density by Year Season for age: ", age, sep="")) +
    theme_bw() + 
    # Customizing facet labels
    theme(strip.text = element_text(size = 12, face = "bold"),
          axis.text = element_text(size = 10), 
          axis.title = element_text(size = 12, face = "bold"))

  ggsave2(filename=paste("age-", age, "_Dens.den.map.jpeg",sep=''),
          ageDens.den.map, device="jpeg", 
          path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
          unit="cm", limitsize = FALSE)

  # Log transformed map
  ageDens.den.tr.map <- 
    ggplot() +
    geom_point(data = ageDens, aes(x=X,y=Y, col = Density.tr)) +
    scale_color_viridis_c() +
    facet_wrap(~YearSeas) +
    labs(title = paste("Density (ln(x+1)) by Year Season for age: ", age, sep="")) +
    theme_bw() +  
    # Customizing facet labels
    theme(strip.text = element_text(size = 12, face = "bold"),
          axis.text = element_text(size = 10), 
          axis.title = element_text(size = 12, face = "bold"))

  ggsave2(filename=paste("age-", age, "_Dens.den.tr.map.jpeg",sep=''),
          ageDens.den.tr.map, device="jpeg", 
          path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
          unit="cm", limitsize = FALSE)
  
  #-----------------------------------------------------------------------------
  # Histogram figures
  #-----------------------------------------------------------------------------

  # For base histogram
  ageDens.pres.den.hist <-
    ggplot() +
    geom_histogram(data = ageDens.pres.set,
                   aes(x = Density, y = ..density..), 
                   bins = 50,  # You can adjust the number of bins or use binwidth
                   fill = "lightblue", color = "black", alpha = 0.7) + 
    labs(title = paste("Histogram of Density by Year Season for age: ", age, sep=""), 
         x = "Density", y = "Density (f(x))") +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_bw() +  
    # Customizing facet labels
    theme(strip.text = element_text(size = 12, face = "bold"),
          axis.text = element_text(size = 10), 
          axis.title = element_text(size = 12, face = "bold"))

  ggsave2(filename=paste("age-", age, "_Dens.pres.den.hist.jpeg",sep=''),
          ageDens.pres.den.hist, device="jpeg", 
          path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
          unit="cm", limitsize = FALSE)

  # For log transformed histogram
  ageDens.pres.den.tr.hist <-
    ggplot() +
    geom_histogram(data = ageDens.pres.set,
                   aes(x = Density.tr, y = ..density..), 
                   bins = 50,  # You can adjust the number of bins or use binwidth
                   fill = "lightblue", color = "black", alpha = 0.7) + 
    labs(title = paste("Histogram of Density by Year Season for age: ", age, sep=""),
         x = "Density (ln(x+1))", y = "Density (f(ln(x+1)))") +
    scale_y_continuous(expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    theme_bw() +  
    # Customizing facet labels
    theme(strip.text = element_text(size = 12, face = "bold"),
          axis.text = element_text(size = 10), 
          axis.title = element_text(size = 12, face = "bold"))

  ggsave2(filename=paste("age-", age, "_Dens.pres.den.tr.hist.jpeg",sep=''),
          ageDens.pres.den.tr.hist, device="jpeg", 
          path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
          unit="cm", limitsize = FALSE)

  #-----------------------------------------------------------------------------
  # Boxplot figs
  #-----------------------------------------------------------------------------

  # For base boxplot
  ageDens.pres.den.boxplot <-
    ggplot() +
    geom_boxplot(data = ageDens.pres.set, 
                 aes(x = YearSeas, y = Density),
                 fill = "lightblue", 
                 color = "black", alpha = 0.7) +
    facet_wrap(~age) %>%
    labs(title = paste("Density by Year Season for age: ", age, sep=""), 
         x = "Year Season", y = "Density (x)") +
    theme_bw() +
    theme(strip.text = element_text(size = 12, face = "bold"),
          axis.text.y = element_text(size = 10),
          axis.text.x = element_text(size = 10, angle = 60, vjust = .5, hjust=.50),
          axis.title = element_text(size = 12, face = "bold"))

  ggsave2(filename= paste("age-", age, "_Dens.pres.den.boxplot.jpeg",sep=''),
          ageDens.pres.den.boxplot, device="jpeg", 
          path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
          unit="cm", limitsize = FALSE)

  # For log transformed boxplot
  ageDens.pres.den.tr.boxplot <-
    ggplot() +
    geom_boxplot(data = ageDens.pres.set, 
                 aes(x = YearSeas, y = Density.tr),
                 fill = "lightblue", 
                 color = "black", alpha = 0.7) +
    facet_wrap(~age) %>%
    labs(title = paste("Density by Year Season for age: ", age, sep=""), 
         x = "Year Season", y = "Density (x)") +
    theme_bw() +
    theme(strip.text = element_text(size = 12, face = "bold"),
          axis.text.y = element_text(size = 10),
          axis.text.x = element_text(size = 10, angle = 60, vjust = .5, hjust=.50),
          axis.title = element_text(size = 12, face = "bold"))

  ggsave2(filename=paste("age-", age, "_Dens.pres.den.tr.boxplot.jpeg",sep=''),
          ageDens.pres.den.tr.boxplot, device="jpeg", 
          path=paste(dir, "/Figures",sep=""), dpi=1200, width = 29, height=21, 
          unit="cm", limitsize = FALSE)
  }

```

# Test/Train Split With Season
 
Split 28 quarters (16 years) for training and 10 quarters (5 years) for out-of-model testing. For the time_series_split function, you need to have a datetime object. 

```{r, split.data2}

tims2 <- data.frame("YearSeas" = sort(unique(TwoSeas.dat$YearSeas))) %>%
  arrange(YearSeas)  # ensures sorted order

# Single split: 75% train, last 3 time points as test
split <- timetk::time_series_split(
  tims2,
  date_var = YearSeas,
  initial = floor(0.75 * nrow(tims2)),
  assess = ceiling(0.25 * nrow(tims2))
)

# Create training set
train_dates2 <- rsample::training(split)
train2 <- subset(TwoSeas.dat, YearSeas %in% train_dates2$YearSeas)
tail(train2$YearSeas)

fwrite(train2, here(dir, "Data", "ExampleRun", "train2"))

# Create test set
test_dates2  <- rsample::testing(split)
test2 <- subset(TwoSeas.dat, YearSeas %in% test_dates2$YearSeas) 
head(test2$YearSeas)

fwrite(test2, here(dir, "Data", "ExampleRun", "test2"))

# Define extra time steps to estimate
est.missing.time2 <- unique(c(missing_seas$YearSeas.nm, test2$YearSeas.nm)) %>%
  sort

save(est.missing.time2, file = here(dir, "Data", "ExampleRun",
                                    "est.missing.time2.rda"))

```

# Construct the Seasonal Mesh: 

see https://pbs-assess.github.io/sdmTMB/articles/basic-intro.html: cutoff is in the units of X and Y (km here), represents minimum distance between knots before a new mesh vertex is added). Muhling et al., 2025 tried a bunch of cutoff values, and chose one that 1) had good out-of-sample predictability, and 2) wasn't too "blocky". They found Values between ~ 50 and 200 were fairly reasonable. 

Here we look at the distance between clusters to assign knots using the summary function.

Important: cannot use anisotropy with barrier mesh (see warning when run sdmTMB).

## Extract and Save Land Barrier Shapefile

```{r, construct.mesh}

# Read full resolution - Continental land masses and ocean islands, except Antarctica.
land.barrier <- read_sf(here(dir, "Data", "Shapefiles", "gshhg-shp-2.3.7",
                             "GSHHS_shp", "f", "GSHHS_f_L1.shp")) %>%
  # Project to same coordinate system
  st_transform(crs = st_crs(area2poly.dens)) %>%
  # Trim to area extent
  st_crop(st_bbox(area2poly.dens))

plot(land.barrier)

# Filter geometry type
land.barrier <- land.barrier[st_geometry_type(land.barrier) %in%
                               c("POLYGON", "MULTIPOLYGON"), ] %>%
  st_union()

# Wrap back into sf object
land.barrier <- st_sf(geometry = land.barrier)

# Check that union worked
st_is_valid(land.barrier)
# [1] TRUE

# Drop levels
land.barrier <- land.barrier %>% mutate(across(where(is.factor), droplevels))

st_write(land.barrier, here(dir, "Data", "Shapefiles", "land.barrier.shp"))

```

## Apply Training and Land Barrier Data and Construct

```{r, construct.seasonal.mesh}

# Cutoff values control the minimum distance between knots.
## It is generally better to start with a coarser mesh (larger cutoff)
## However there is a tradeoff on spatial predictability (more knots) and
## over fitting the time to process. If the day is irregularly distributed
## you can also try residual-based knot placement
meshTrain2 <- make_mesh(train2, xy_cols = c("X", "Y"), cutoff = 50) 

# Check number of mesh nodes 
## If greater than >1,000–2,000 nodes, 
## you’re likely in trouble unless you have a lot of RAM.
length(meshTrain2$mesh$loc[,1])
# [1] 212

# proj_scaling should match units of (since we are working in m, but density
# is in km, divide by 1000)
barrier2 <- add_barrier_mesh(meshTrain2, land.barrier, proj_scaling = 1000)

## If greater than >1,000–2,000 nodes, 
## you’re likely in trouble unless you have a lot of RAM.
length(barrier2$mesh$loc[,1])
# [1] 604

# Example plot code from ?add_barrier_mesh
mesh_df_water2 <- barrier2$mesh_sf[barrier2$normal_triangles, ]
mesh_df_land2 <- barrier2$mesh_sf[barrier2$barrier_triangles, ]

Mesh.with.barrier2 <-
ggplot() +
  geom_sf(data = land.barrier) +
  geom_sf(data = mesh_df_water2, size = 1, colour = "blue") +
  geom_sf(data = mesh_df_land2, size = 1, colour = "green")

Mesh.with.barrier2
# [1] 231

plot(barrier2)
points(train2$X, train2$Y, col = "red", pch = 19, 
       cex = 0.2)

# Define extra time steps to estimate
est.missing.time2 <- unique(c(missing_seas$YearSeas.nm, test2$YearSeas.nm)) %>%
  sort

```

